---
title: "Pilot data exploration"
author: "MW"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

# Load required libraries

```{r setup, include=FALSE}

# Ensure the necessary libraries are loaded
library(here, warn.conflicts = FALSE, quietly=TRUE)
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
library(RColorBrewer)
# Interpolate missing values using linear interpolation for both eyes
library(zoo)

# set working directory
data_dir <- paste0(here())
setwd(data_dir)
```

# Reading files

**Function for generating mock datasets**

```{r }
# Create a helper function to handle NA interpolation with leading and trailing NAs
na_approx_with_padding <- function(x) {
  # Interpolate only non-NA segments
  x_approx <- zoo::na.approx(x, na.rm = FALSE)
  
  # Replace leading and trailing NAs with original values
  x_approx[is.na(x_approx)] <- x[is.na(x_approx)]
  
  return(x_approx)
}

```

**Reading the device and question data from all participants**

### Answers analysis

**Distribution of answers**

```{r }

# Initialize lists to store data
deviceDataList <- list()
questionDataList <- list()

# Number of participants
participant_ids <- c(44)

# Loop through each participant
for (id in participant_ids) {
  # Generate the file names dynamically
  deviceFileName <- paste0("data/", id, "_DeviceData.csv")
  questionFileName <- paste0("data/", id, "_QuestionsData.csv")
  
  # Read the CSV files
  deviceData <- read.csv(deviceFileName, sep = ";", encoding = "utf-8")
  questionData <- read.csv(questionFileName, sep = ";", encoding = "utf-8")
  
  questionData["Id"] <- id
  # Append the data to the lists
  if (!is.null(deviceData)) deviceDataList[[length(deviceDataList) + 1]] <- deviceData
  if (!is.null(questionData)) questionDataList[[length(questionDataList) + 1]] <- questionData
    
}

questionData_combined <- bind_rows(questionDataList)

summary(questionData_combined$AnswerDuration)

# Summarize the data
answers_summary <- questionData_combined %>%
  group_by(Id, VideoMode, VideoName, QuestionIndex, QuestionAnswer) %>%
  summarize(Count = n(), .groups = 'drop')

# Display the summary table
kable(answers_summary[1:10,], caption = "Summary Table of Question Answers by Participant, VideoMode, and VideoName")

# Summarize the average QuestionAnswer for each VideoMode and VideoName

summary_data <- questionData_combined %>%
  group_by(VideoName, VideoMode) %>%
  summarize(
    Mean_Answer = mean(QuestionAnswer, na.rm = TRUE),
    SD_Answer = sd(QuestionAnswer, na.rm = TRUE),
    Mean_Duration = mean(AnswerDuration, na.rm = TRUE),
    SD_Duration = sd(AnswerDuration, na.rm = TRUE),
    .groups = "drop"
  )

# Display the summarized data
kable(summary_data, caption = "Average QuestionAnswer for each  VideoName")

# Define custom colors for the categories
custom_colors <- c(
  "Hallucination" = "#4A1B75", # RGB(74, 27, 117)
  "Coupling" = "#1B7A80",      # RGB(27, 122, 128)
  "Normal" = "#464640"         # RGB(70, 70, 64)
)

ggplot(summary_data, aes(x = VideoMode, y = Mean_Answer)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  geom_errorbar(aes(ymin = Mean_Answer - SD_Answer, ymax = Mean_Answer + SD_Answer), 
                position = position_dodge(width = 0.8), width = 0.2) +
  scale_fill_manual(values = custom_colors) + # Apply custom colors
  theme_minimal() +
  labs(
    title = "Mean Question Answer by VideoMode",
    x = "Video Mode",
    y = "Mean Question Answer"
  )


ggplot(questionData_combined, aes(x = VideoMode, y = QuestionAnswer)) +
  geom_boxplot() +
  scale_fill_manual(values = custom_colors) + # Apply custom colors
  theme_minimal() +
  labs(
    title = "Distribution of Question Answers by VideoMode",
    x = "Video Mode",
    y = "Question Answer"
  )


```

**Statistical tests of answers**

```{r }

# ANOVA to compare QuestionAnswers across different VideoModes and VideoNames
anova_result <- aov(QuestionAnswer ~ VideoMode * VideoName, data = answers_summary)
summary(anova_result)


```

### Eye tracking

## 44

```{r}
deviceData <- deviceDataList[[1]]
deviceData <- deviceData %>% select_if(~ !any(is.na(.)))

### Rename some columns to fit the code from VR_ET_code.Rmd (sent by PM to MW)
deviceData <- deviceData %>%
  rename(
    HMDPosition = DevicePosition,
    HMDRotation = DeviceRotation,
    LeftEyeOpenness = LeftEyeOpeness,
    RightEyeOpenness = RightEyeOpeness,
    systemTime = ExperimentSecond

  )

deviceData <- deviceData %>% select_if(~ !any(is.na(.)))
deviceData <- deviceData[-1, ]

deviceData$RightEyePupilSize[deviceData$RightEyePupilSize == -1] <- NA
deviceData$LeftEyePupilSize[deviceData$LeftEyePupilSize == -1] <- NA

```

### Helper functions for time markers :

##### Event changed based on ECG data

```{r}

# df should be ECG data

when_event_changed <- function(df) {
    result <- df %>%
    mutate(video_start = (EVENT == 4 & lag(EVENT) != 4)) %>%
    mutate(video_end = (EVENT == 8 & lag(EVENT) != 8)) %>%
    filter(video_start == TRUE |video_end==TRUE)
  
  video_start <- result %>%
    filter(video_start==TRUE)
  
  
  video_end <- result %>%
    filter(video_end==TRUE)
  
  video_start <- video_start$systemTime
  video_end <- video_end$systemTime
  
  print(video_start)  ## this is corresponds to segment change (when new video starts)
  print(video_end)
  
  return(result)
  
}
when_event_changed(ecgData_cleaned)

```

##### Segment change based on deviceData

```{r}
# df should be device data
when_segment_changed <- function(df) {
  result <- df %>%
    mutate(segment_changed = (SegmentIndex != lag(SegmentIndex))) %>%
    filter(segment_changed == TRUE)
  
  return(result$systemTime)
}
when_segment_changed(deviceData)
```

##### Movie mode changed based on Device Data

```{r}
# Function that takes deviceData df and returns a frame with rows, when the VideoMode/Video changed 
when_movie_mode_changed <-function(df) {
  # Find changes in VideoName
  name_changes <- c(FALSE, df$VideoName[-1] != df$VideoName[-nrow(df)])
  # Find changes in VideoMode
  mode_changes <- c(FALSE, df$VideoMode[-1] != df$VideoMode[-nrow(df)])
  # Combine the two logical vectors
  changes <- name_changes | mode_changes
  
  # Get the indices where changes occur
  change_indices <- which(changes)
  
  changes_rows <- df[change_indices,]
  return(changes_rows)
}

when_movie_mode_changed(deviceData)$systemTime
```

```{r}
when_movie_mode_changed_ext <- function(df) {
  # Ensure df has the necessary columns
  if (!all(c("systemTime", "VideoName", "VideoMode") %in% colnames(df))) {
    stop("Data frame must contain 'systemTime', 'VideoName', and 'VideoMode' columns.")
  }
  
  # Find changes in VideoName
  name_changes <- c(FALSE, df$VideoName[-1] != df$VideoName[-nrow(df)])
  # Find changes in VideoMode
  mode_changes <- c(FALSE, df$VideoMode[-1] != df$VideoMode[-nrow(df)])
  
  # Get the indices where changes occur
  name_change_indices <- which(name_changes)
  mode_change_indices <- which(mode_changes)
  
  print(name_change_indices)
  print(mode_change_indices)
  
  # Create data for VideoName changes
  name_changes_rows <- data.frame(
    systemTime = df$systemTime[name_change_indices],
    type = "moviename",
    name = df$VideoName[name_change_indices],
    mode = df$VideoMode[name_change_indices],

    stringsAsFactors = FALSE
  )
  

  
  return(name_changes_rows)
}

```

##### Movie/mode changed based on DeviceData (VideoName, VideoMode columns)

```{r}
movie_mode_change = when_movie_mode_changed(deviceData)[["systemTime"]]
movie_mode_change_ext = when_movie_mode_changed_ext(deviceData)
```

##### Segment changed based on DeviceData (Segment Column)

```{r}
segment_changed <- when_segment_changed(deviceData)
segment_changed
```

##### Event changed based on ECG data

```{r}
ecgData <- read.csv("data/44_ECG.csv", sep = ",", encoding = "utf-8")

ecgData <- clean_and_add_time_to_ecg(ecgData,deviceData$systemTime[1])

event_changed <- when_event_changed(ecgData)
```

##### Comparing all changes (ecg, ET, segment, etc.)

```{r}
video_start_ecg <- (event_changed %>% filter(video_start))$systemTime

video_end_ecg <- (event_changed %>% filter(video_end))$systemTime



# Create data frames for each event
df_segment <- data.frame(time = segment_changed, event = 'Segment Change(deviceData)')
df_video_start <- data.frame(time = video_start_ecg, event = 'Video Start (ECG)')
df_video_end <- data.frame(time = video_end_ecg, event = 'Video End (ECG)')
df_movie_mode <- data.frame(time = movie_mode_change, event = 'Movie/Mode Change(deviceData)')

# Combine the data frames into one
changes_combined <- rbind(df_segment, df_video_start, df_video_end, df_movie_mode)

changes_combined <- changes_combined %>%
  arrange(time)
```

##### Plotting the time events (all sources)

```{r}

ggplot(changes_combined, aes(x = time, y = event, color = event)) +
  geom_point(size = 3) +
  geom_vline(xintercept = segment_changed, linetype = "dashed", color = "blue", alpha = 0.5) +
  geom_vline(xintercept = video_start_ecg, linetype = "dotted", color = "green", alpha = 0.5) +
  geom_vline(xintercept = video_end_ecg, linetype = "solid", color = "red", alpha = 0.5) +
  geom_vline(xintercept = movie_mode_change, linetype = "longdash", color = "purple", alpha = 0.5) +
  labs(title = "Time Event Plot", x = "Time (seconds)", y = "Event Type") +
  theme_minimal() +
  scale_color_manual(values = c("Segment Change(deviceData)" = "blue", "Video Start (ECG)" = "green", 
                                "Video End (ECG)" = "red", "Movie/Mode Change(deviceData)" = "purple")) +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
```

## Plotting events from ECG

```{r}
ecg_no_zero <- ecgData %>% filter(EVENT >0)
pdf("EVENT_vs_systemTime.pdf", width = 28, height = 16)  # Adjust width and height as needed
plot(ecg_no_zero$systemTime, ecg_no_zero$EVENT, xlab = "systemTime", ylab = "EVENT", 
     main = "EVENT vs systemTime", pch = 1, col = "blue")

dev.off()
```

##### Smoothed Pupile Size (approx. function)

```{r}

# Apply approx function to the pupil size columns
deviceData <- deviceData %>%
  arrange(systemTime) %>%
  mutate(
    RightEyePupilSize = na_approx_with_padding(RightEyePupilSize),
    LeftEyePupilSize = na_approx_with_padding(LeftEyePupilSize),
    smoothed_right_pupil_size = zoo::rollmean(RightEyePupilSize, k = 180, fill = NA),
    smoothed_left_pupil_size = zoo::rollmean(LeftEyePupilSize, k = 180, fill = NA)
  )

# Create a line plot with smoothing for both eyes
ggplot(deviceData, aes(x = systemTime)) +
  geom_line(aes(y = smoothed_right_pupil_size), color = "blue", na.rm = TRUE) +
  geom_line(aes(y = smoothed_left_pupil_size), color = "red", na.rm = TRUE) +
  labs(title = "Smoothed Pupil Size Over Time",
       x = "Relative Time", y = "Smoothed Pupil Size",
       color = "Eye") +
  scale_color_manual(values = c("Right Eye" = "blue", "Left Eye" = "red"))+
  geom_vline(xintercept = video_start_ecg, color = "green", size = 0.5, linetype = 4)+
  geom_vline(xintercept = video_end_ecg, color = "pink", size = 0.5, linetype = 4)


```

##### Average Pupil Size

```{r}
# Calculate the average pupil size
deviceData$average_pupil_size <- rowMeans(deviceData[, c("smoothed_right_pupil_size", "smoothed_left_pupil_size")], na.rm = TRUE)


# Create a line plot for the average pupil size
# removed vlines as no spawnData was identified
ggplot(deviceData, aes(x = systemTime, y = average_pupil_size)) +
  geom_line(color = "purple") +
  labs(title = "Average Pupil Size Over Time",
       x = "Relative Time", y = "Average Pupil Size")+
 geom_vline(xintercept = video_start_ecg, color = "darkgreen", size = 0.5, linetype = 4)+
  geom_vline(xintercept = video_end_ecg, color = "red", size = 0.5, linetype = 4)
```

```{r}
video_start_ecg
```

```{r}
video_end_ecg
```

```{r}
split_device_data <- list() # List to store split dataframes
# Loop through each pair of start and end times
for (i in seq_along(video_start_ecg)) {
  # Filter the dataframe for rows within the start and end times
  split_device_data[[i]] <- deviceData %>%
    filter(systemTime >= video_start_ecg[i] & systemTime <= video_end_ecg[i])
}

# Optionally, assign names to each dataframe in the list
names(split_device_data) <- paste0("segment_", seq_along(video_start_ecg))

# Example: Access the first split dataframe
segment_1 <- split_device_data[[1]]
```

```{r}

# Adjust `systemTime` to start from 0 for each segment
combined_device_data <- bind_rows(
  lapply(seq_along(split_device_data), function(i) {
    segment <- split_device_data[[i]]
    segment$segment <- paste0("Segment ", i) # Add segment identifier
    segment$systemTime <- segment$systemTime - video_start_ecg[i] # Adjust systemTime
    return(segment)
  })
)


# Create the plot
ggplot(combined_device_data, aes(x = systemTime)) +
  geom_line(aes(y = smoothed_right_pupil_size, color = "Right Eye"), na.rm = TRUE) +
  geom_line(aes(y = smoothed_left_pupil_size, color = "Left Eye"), na.rm = TRUE) +
  labs(
    title = "Smoothed Pupil Size Over Time for All Segments (Adjusted System Time)",
    x = "Adjusted System Time",
    y = "Smoothed Pupil Size",
    color = "Eye",
  ) +
  scale_color_manual(values = c("Right Eye" = "blue", "Left Eye" = "red")) +
  facet_wrap(~segment, scales = "free_x") +
  theme_minimal()


```

```{r}
# Create the plot for average pupil size
ggplot(combined_device_data, aes(x = systemTime, y = average_pupil_size)) +
  geom_line(color = "purple") +
  labs(
    title = "Average Pupil Size Over Time for All Segments (Adjusted System Time)",
    x = "Adjusted System Time",
    y = "Average Pupil Size"
  ) +
  facet_wrap(~segment, scales = "free_x") + # Separate plots for each segment
  theme_minimal()
```

```{r}
plot(head(ecgData$ECG100,2500), type = "l")
```
